# Project : Reproducible Research


## 1- Getting the data
Read the data.   
```{r}
data     <- read.csv("activity.csv", stringsAsFactors=FALSE)
```

## 2- What is mean total number of steps taken per day?
Ignore the missing values in the dataset, and Calculate the total number of steps taken per day.   
We also calculate and report the mean and median of the total number of steps taken per day.

```{r}
library(plyr)
data1 <- ddply(data,.(date),summarize,totalsteps=sum(steps, na.rm = TRUE))
hist(data1$totalsteps,col="pink",breaks=40,main="Total steps / day",xlab="Steps /day")
abline(v=mean(data1$totalsteps),lty=3, col="red")
abline(v=median(data1$totalsteps),lty=2, col="blue")
legend("topright", legend=c("mean", "median"),  col=c("red", "blue"), lty=c(3,2), cex=0.8)
```

## 3- What is the average daily activity pattern?
Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
data2 <-ddply(data,.(interval),summarize,meansteps=mean(steps, na.rm = TRUE))
plot(data2$interval, data2$meansteps,type="l",col="red",main="average daily activity", xlab="intervals", ylab="average")
abline(v=data2$interval[which.max(data2$meansteps)], lty=2, col="blue")
legend("topright", legend=" the interval with the average of the maximum number of steps",  col="blue", lty=2, cex=0.8)
```

## 4- Imputing missing values
Calculate and report the total number of missing values in the dataset
```{r}
number_na <- sum(is.na(data$steps))
number_na
```

Create a new dataset that is equal to the original dataset but with the missing data filled in by using the average number of steps in particular 5-minutes interval
```{r}
data3 <- data
for (i in 1:nrow(data3))
{  if (is.na(data3[i,"steps"]))
	   data3[i,"newsteps"] <- subset(data2$meansteps, (data2$interval == data3[i,"interval"]))
   else
	   data3[i,"newsteps"] <- data3[i,"steps"]
}
```

look what new data like
```{r}
head(data3)
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r}
data4 <- ddply(data3,.(date),summarize,totalsteps=sum(steps, na.rm = TRUE))
hist(data4$totalsteps,col="green",breaks=40,main="Total steps / day",xlab="Steps /day")
abline(v=mean(data4$totalsteps),lty=3, col="red")
abline(v=median(data4$totalsteps),lty=2, col="blue")
legend("topright", legend=c("mean", "median"),  col=c("red", "blue"), lty=c(3,2), cex=0.8)	
```

## 5-Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
library(lubridate)
for (i in 1:nrow(data3))
{  if (wday(as.Date(data3[i,"date"]) ) %in%c(1,7) )
	   data3[i,"newday"] <- "weekend"
   else
	   data3[i,"newday"] <- "weekday"
}
data3$newday <- as.factor(data3$newday)
```

Summary of the new variable
```{r}
table(data3$newday)
```

Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r}
library(lattice)
xyplot(newsteps~interval|newday,data=data3,type='a',layout=c(1,2))
```